---
title: "Pre Processing for VaR"
output: github_document
authors: "Jack Parkin, Sihguat Torres, Shaan Patel, Isaac Arocha"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---


Run this program with the period specification set at 5 and save raw_data_list as some variable! Then re-run the program with the period specification at 25, and you'll have all of the data in two lists - one for each period specification.

Period Specification
```{r warning=FALSE, message=FALSE}
# 1 Year Scenario -> we run from 1990 to 2014, total of 25 years including 1990. i = 1:25
# 5 year scenario -> we run from 1990 to 2014, total of five 5-year periods. i = 1:5

periods <- 5 #CAN ONLY TAKE ON THE VALUE OF 5 OR 25
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries
```{r warning=FALSE, message=FALSE}
# Used Throughout
library(lubridate)
library(quantmod)
library(tidyverse)
library(data.table)
library(readxl) 
library(here) # Instead of setting WD with setwd, use the here() function for code transferability between devices

# # For Logistic Model
# library(leaps)
# library(ROCR)
# library(glmnet)
# 
# # For Neural Net Model
# library(neuralnet)

#Extra
library(devtools)
```



# Data Import 
```{r warning=FALSE, message=FALSE}
raw_data <- read.csv(here("input_pre_processing_VAR.csv"))
```

# Pre-Processing

The goal here is to split our timeline (1990-2014) into periods, get a dataframe for each period, and remove the inactive loans from each period.
Active here is defined as the following: the loan was approved in or before that period AND the loan was terminated in or after that period.

First, I find the terminus of each loan. 

```{r}
raw_data <- 
  raw_data %>% 
  mutate(
    term_years = TermInMonths %/% 12,
    terminus = 
      case_when(
        !is.na(ChargeOffDate) ~ as.integer(substring(ChargeOffDate, 0, 4)), # Loan is charged off
        ApprovalFiscalYear + term_years <= 2013 ~ as.integer(ApprovalFiscalYear + term_years),  # Loan is paid in full prior to end of dataset
        ApprovalFiscalYear + term_years > 2013 ~ as.integer(2013),
        TRUE ~ NA_integer_
      )
  ) 
```

Now, I copy the raw_data set as many times as there are periods in our data set. 
```{r}
raw_data_list <- list()

for (i in 1:periods) { 
  raw_data_list[[i]] = raw_data
}
```

Now that I've got my list of data sets, I'm going to walk through that list and eliminate the inactive loans within each of the list's elements.

```{r}
for (i in 1:periods) {
  
  if (periods == 5) {
    active_years = seq(1990 + 5*(i - 1), 1990 + 5*i, 1)
    raw_data_list[[i]] <- 
    raw_data_list[[i]] %>% 
    filter(
      ApprovalFiscalYear %in% active_years | ApprovalFiscalYear < active_years[1],
      terminus %in% active_years | terminus > active_years[6]
    )
  } else {
    curr_year = 1990 + i - 1
    raw_data_list[[i]] <- 
    raw_data_list[[i]] %>% 
    filter(
      ApprovalFiscalYear <= curr_year,
      curr_year <= terminus
    )
  }
  
}
```

At this point, I've got a list of data frames, one per period. Each data frame consists only of the active loans during that period. 
Now, I've got to add in the other additional variables to each data set.


# Additional Data
```{r warning=FALSE, message=FALSE}
#GDP Import
GDP = read.csv("GDP.csv")

#SP500 Import
SP500 = getSymbols("^GSPC",from = "1990-01-31",to = "2014-12-31", periodicity = 'monthly', auto.assign = FALSE)

#FedFunds Import
FEDFUNDS = read.csv("FEDFUNDS.csv")

#CPI Import (Consumer Price Index for All Urban Consumers: All Items in U.S. City Average)
CPI = read.csv("CPIAUCSL.csv")

#Unemployment
UnemploymentUSbyState = read.csv('StateUR.csv', header = TRUE)
```

Prepping the GDP Data

```{r}
GDP <- 
  GDP %>%  
  transmute(month.bin = DATE, GDP = GDP) %>%  
  transmute(
    GDP,
    quarter.bin =
      case_when(
        as.numeric(substring(month.bin, 6, 7)) < 10 & as.numeric(substring(month.bin, 6, 7)) >= 7  ~ paste("Q3", substring(month.bin, 0, 4)),
        as.numeric(substring(month.bin, 6, 7)) < 7 & as.numeric(substring(month.bin, 6, 7)) >= 4   ~ paste("Q2", substring(month.bin, 0, 4)),
        as.numeric(substring(month.bin, 6, 7)) < 4                                                 ~ paste("Q1", substring(month.bin, 0, 4)),
        TRUE                                                                                       ~ paste("Q4", substring(month.bin, 0, 4)),
      )
  )
```

Prepping the S&P 500 Data

```{r}
SP500 <- 
  SP500 %>% 
  as.data.frame() %>% 
  transmute(month.bin = rownames(.), GSPC.price = GSPC.Adjusted)
```

Adding all of the additional, time-varying data to each item in the list of data sets

```{r}

for (i in 1:periods) {
  repeat_borr = duplicated(raw_data_list[[i]]$BorrName)
  
  raw_data_list[[i]] <- 
    raw_data_list[[i]] %>%  
    mutate(
      quarter.bin =
        case_when(
          as.numeric(substring(ApprovalDate, 6, 7)) < 10 & as.numeric(substring(ApprovalDate, 6, 7)) >= 7  ~ paste("Q3", substring(ApprovalDate, 0, 4)),
          as.numeric(substring(ApprovalDate, 6, 7)) < 7 & as.numeric(substring(ApprovalDate, 6, 7)) >= 4   ~ paste("Q2", substring(ApprovalDate, 0, 4)),
          as.numeric(substring(ApprovalDate, 6, 7)) < 4                                                    ~ paste("Q1", substring(ApprovalDate, 0, 4)),
          TRUE                                                                                             ~ paste("Q4", substring(ApprovalDate, 0, 4)),
        )
    ) %>%  
    left_join(y = GDP, by = "quarter.bin") %>%  
    subset(select = -quarter.bin) %>%     # Up to here, I've added GDP. Onto S&P 500
    left_join(y = SP500, by = "month.bin") %>%  
    left_join(y = FEDFUNDS, by = c("month.bin" = "DATE")) %>%  
    left_join(y = CPI, by = c("month.bin" = "DATE")) %>% 
    left_join(y = UnemploymentUSbyState, by = c("ProjectState" = "State", "month.bin" = "DATE")) %>%  
    rename(URinBorrState = UR) %>%  
    left_join(y = UnemploymentUSbyState, by = c("BorrState" = "State", "month.bin" = "DATE")) %>%  
    rename(URinProjectState = UR) %>% 
    mutate(
      BinaryIntergerTerm = case_when(TermInMonths %% 12 == 0 ~ 1L, TRUE ~ 0L),
      BinaryThirdPartyDollars = 
        case_when(
          ThirdPartyDollars ~ 1L, 
          !ThirdPartyDollars ~ 0L, 
          TRUE ~ NA_integer_
        ),
      BinaryBankStEqualBorrowerSt = 
        case_when(
          as.character(BorrState) == as.character(CDC_State) ~ 1L, 
          as.character(BorrState) != as.character(CDC_State)~ 0L,
          TRUE ~ NA_integer_
      ),
      BinaryProjectStEqualBorrowerSt =
        case_when(
          as.character(BorrState) == as.character(ProjectState) ~ 1L, 
          as.character(BorrState) != as.character(ProjectState) ~ 0L, 
          TRUE ~ NA_integer_
      ),
      BinaryRepeatBorrower = case_when(
          repeat_borr ~ 1L, 
          !repeat_borr ~ 0L, 
          TRUE ~ NA_integer_
      ),
      Default = 
        case_when(
          GrossChargeOffAmount == 0 ~ 0L,
          TRUE ~ 1L
      ),
      LogGrossApproval = log(GrossApproval),
      LogGDP = log(GDP),
      LogSP500 = log(GSPC.price),
      LogFedFunds = log(FEDFUNDS),
      LogCPI = log(CPIAUCSL)
    )    
    
}
```


```{r}
for (i in 1:periods) {
  raw_data_list[[i]] <-
    raw_data_list[[i]] %>% 
    mutate(
      across(
        .cols =
          c(
            LogGrossApproval,
            TermInMonths,
            LogGDP,
            LogSP500,
            LogFedFunds,
            LogCPI,
            URinProjectState,
            URinBorrState
          ),
        .fns = ~ (. - mean(., na.rm = TRUE)) / (sd(., na.rm = TRUE))
      )
    )
  
  raw_data_list[[i]] <- 
    raw_data_list[[i]][c("GrossApproval", "GrossChargeOffAmount", "Default", "LogGrossApproval", "BinaryThirdPartyDollars", "TermInMonths", "LogGDP", "LogSP500", "LogFedFunds", "LogCPI", "URinProjectState", "URinBorrState", "BusinessType", "NAICS_Sector", "DeliveryMethod", "BinaryIntergerTerm", "BinaryRepeatBorrower", "BinaryBankStEqualBorrowerSt", "BinaryProjectStEqualBorrowerSt", "ApprovalFiscalYear", "BorrState", "ProjectState")]
  
  #Dealing with missing values!
  raw_data_list[[i]] <- 
    raw_data_list[[i]] %>% 
      mutate(
        across(
          .cols = 
            c(
              LogGrossApproval,
              TermInMonths,
              LogGDP,
              LogSP500,
              LogFedFunds,
              LogCPI,
              URinProjectState,
              URinBorrState
            ),
          .fns = ~ if_else(is.na(.), 0, .)
        ),
        across(
          .cols = 
            c(
              BusinessType, 
              Default,
              NAICS_Sector, 
              DeliveryMethod, 
              BinaryIntergerTerm, 
              BinaryRepeatBorrower, 
              BinaryBankStEqualBorrowerSt, 
              BinaryProjectStEqualBorrowerSt, 
              ApprovalFiscalYear, 
              BorrState, 
              ProjectState, 
              BinaryThirdPartyDollars
            ),
          .fns = ~ if_else(is.na(.), "Blank", as.character(.))
      )
  )
}
```


Exporting the Data
```{r}
for (i in 1:periods) {
  
  file_name <- 
    if_else(
      periods == 5,
      paste("5year_VARdata_period", "_", as.character(i), ".csv", sep = ""),
      paste("1year_VARdata_period", "_", as.character(i), ".csv", sep = "")
    )
  
  write_csv(raw_data_list[[i]], path = here(file_name))
}
```


